\section*{Overview}

Knuth optimization is the interval-DP speedup for recurrences of the form
\[
dp[l][r] = \min_{k \in [l, r-1]} \bigl(dp[l][k] + dp[k+1][r]\bigr) + C(l, r),
\]
when the best split point moves monotonically with the interval.

\section*{When to Use It}

Use it when:

\begin{itemize}
  \item the DP is on intervals,
  \item the transition tries every split point,
  \item the cost satisfies the quadrangle-inequality style conditions that imply
  \[
  opt[l][r-1] \le opt[l][r] \le opt[l+1][r].
  \]
\end{itemize}

\section*{Core Idea}

Instead of searching all split points for every interval, only search between the previously known neighboring optimal
split points.

\section*{Key Insight}

The speedup is not ``because interval DP is nice.'' It is because the argmin drifts slowly. Once you know the optimum
split for adjacent intervals, the next optimum is trapped between them.

\section*{Worked Problem}

\paragraph{Problem.}
You have adjacent files with sizes \(a_1, \dots, a_n\). Merging two neighboring groups costs the total size of the
merged group. Find the minimum total merge cost.

\paragraph{Why Knuth fits.}
The ordinary interval DP is cubic. This cost function satisfies the needed monotonicity, so the split search range can
be narrowed to the Knuth interval.

\section*{Correctness Intuition}

The theorem behind Knuth optimization guarantees the monotone-opt property. Once that is true, searching outside
\([opt[l][r-1], opt[l+1][r]]\) cannot improve the answer because the argmin for \([l,r]\) is known to stay inside.

\section*{Complexity Analysis}

The complexity drops from \(O(n^3)\) to \(O(n^2)\).

\section*{Implementation}

The sample code solves the adjacent-file merging problem with prefix sums and Knuth's search bounds.

\section*{Common Pitfalls}

\begin{itemize}
  \item Applying Knuth optimization without proving the monotone-opt condition.
  \item Confusing it with divide-and-conquer DP optimization; the recurrences are different.
  \item Filling intervals in the wrong order and reading \texttt{opt} values before they exist.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Optimal BST-style interval DP.
  \item Any interval merge or partition DP with the right quadrangle structure.
  \item Divide-and-conquer optimization when the recurrence is one-sided instead of interval-based.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item File merging with adjacent merges.
  \item Optimal BST-style interval partitioning.
  \item Any interval DP where split-point monotonicity can be proved.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/dynamic_programming/knuth-optimization.html}{cp-algorithms: Knuth's Optimization}.
  \item \href{https://usaco.guide/adv/knuth?lang=cpp}{USACO Guide: Knuth Optimization}.
  \item \href{https://codeforces.com/catalog}{Codeforces Catalog}.
\end{itemize}
