\section*{Overview}

Rollback DSU is the union-find structure for offline problems where merges must be undone later. The central design
choice is deliberate:

\begin{quote}
Give up path compression so every merge changes only a constant amount of reversible state.
\end{quote}

\section*{Problem-Driven Motivation}

The canonical contest setting is dynamic connectivity:

\begin{itemize}
  \item edges are added and removed over time,
  \item connectivity queries ask about specific times,
  \item the online fully dynamic problem is hard, but the full event list is known offline.
\end{itemize}

Ordinary DSU handles edge additions perfectly, but once an edge becomes inactive, ordinary DSU has no clean undo. The
rollback version is built specifically for the offline divide-and-conquer or segment-tree-over-time reduction.

\section*{Recognition Pattern}

Rollback DSU is a strong fit when:

\begin{itemize}
  \item unions must later be undone,
  \item the problem is offline,
  \item recursion branches or time intervals each need their own reversible DSU state.
\end{itemize}

Typical signals are:

\begin{itemize}
  \item offline dynamic connectivity,
  \item ``apply these changes on an interval, recurse, then undo'',
  \item component data that evolves over a segment tree on time.
\end{itemize}

\section*{Derivation}

Ordinary DSU is fast because of path compression, but path compression is also exactly what makes undo hard: it changes
many parent pointers at once.

So rollback DSU uses:

\begin{itemize}
  \item union by size or rank,
  \item no path compression,
  \item a history stack recording the old state of every successful merge.
\end{itemize}

Each successful union changes only:

\begin{itemize}
  \item one child's parent,
  \item one root's size,
  \item optionally the component counter.
\end{itemize}

That is a constant-sized record, so rollback is simply popping history entries until a saved snapshot size is restored.

\section*{Worked Problem}

\paragraph{Problem.}
Edges are active on time intervals. For many time points, answer whether two vertices are connected.

\paragraph{Why naive fails.}
Rebuilding the graph and recomputing connectivity from scratch at every time point is far too slow.

\paragraph{Offline reduction.}
Build a segment tree over time. If edge \(e\) is active on interval \([l,r]\), attach \(e\) to all segment-tree nodes
whose time segments lie fully inside \([l,r]\).

\paragraph{DFS over time.}
At one segment-tree node:

\begin{itemize}
  \item take a snapshot of the DSU history size,
  \item unite all edges stored at that node,
  \item recurse to children,
  \item roll back to the snapshot before returning.
\end{itemize}

\paragraph{Why this works.}
Every recursion path corresponds to one time point. Along that path, exactly the edges active at that time have been
merged into the DSU.

\section*{Implementation Reasoning}

The rollback interface should stay minimal:

\begin{itemize}
  \item \texttt{find} without compression,
  \item \texttt{unite},
  \item \texttt{snapshot},
  \item \texttt{rollback(snapshot)}.
\end{itemize}

The important invariant is:

\begin{quote}
Rolling back to a snapshot must restore the DSU to exactly the state it had when that snapshot was taken.
\end{quote}

That is why every successful merge records the overwritten parent, overwritten size, and component-count effect.

\section*{Correctness Intuition}

On the segment tree over time, entering a node means ``all edges attached here are active for every time in this whole
interval.'' So it is safe to merge them before recursion. Rolling back after finishing that interval ensures sibling
intervals do not inherit merges that should not exist there.

\section*{Complexity Analysis}

With union by size and no path compression:

\begin{itemize}
  \item \texttt{find}: \(O(\log n)\) worst case,
  \item \texttt{unite}: \(O(\log n)\),
  \item rollback of one merge: \(O(1)\),
  \item memory: \(O(n + \text{number of successful merges})\).
\end{itemize}

\section*{Common Pitfalls}

\begin{itemize}
  \item Adding path compression and destroying reversibility.
  \item Forgetting to take the snapshot before applying the interval's merges.
  \item Recording too little history to restore the old state exactly.
  \item Using rollback DSU on an online deletion problem where the full event timeline is not known.
\end{itemize}

\section*{Variants and Failure Modes}

\begin{itemize}
  \item Additional component metadata such as parity or sums can be rolled back too.
  \item Persistent DSU is a different idea: version access instead of explicit undo.
  \item If the problem needs online edge deletions, rollback alone is not enough.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Offline dynamic connectivity.
  \item Offline bipartiteness under interval-active edges.
  \item Divide-and-conquer-over-time problems with reversible component merges.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/data_structures/deleting_in_log_n.html}{cp-algorithms: Deleting from a Data Structure in \(O(T(n)\log n)\)}.
  \item \href{https://github.com/kth-competitive-programming/kactl}{KACTL}.
  \item \href{https://codeforces.com/catalog}{Codeforces Catalog}.
\end{itemize}
